---
title: "Final assignment"
author: "Uine Kailamäki"
date: "2021"
output: html_document
---


# Final assignment

## Task 1

Using Dataset 1, 
a) plot medieval settlements in the county of Kent, and 
b) run K, L and Pair Correlation Function analysis with 99 runs of Monte Carlo simulation on the data. 

```{r}

# Setting up WD and necessary libraries
setwd("~/R/SADE/week7")
library(rgdal)   
library(raster)   
library(spatstat) 
library(maptools)
library(GISTools)  


# Loading and plotting the polygon "kent"
polyg <- readOGR(dsn="kent")
plot(polyg)

# Adding Digital Elevation Model to the plot to visualize topography
dem <- raster("dem_england/dem_england_historic.tif")
plot(dem, add=T, col=terrain.colors(15))

# Adding the medieval settlements as points to the plot
sett <- readOGR(dsn="dataset1")
points(sett, pch=19, cex=0.2)

# Omitting points that might fall outside of the polygon 
sett <- sett[polyg, ]

# Inspecting how "dataset1" looks like
head(sett)
summary(sett)

# Converting the data into spastat object for analysis
sp_sett <- as.ppp(coordinates(sett),as.owin(polyg))

# K-Function analysis
# "Pois" is the expected Poisson line, the others are how the data compares to it with a few different edge corrections
k_fun <- Kest(sp_sett)
plot(k_fun, xlim=c(0,5000), main="K-Function analysis")

# L-Function analysis
# Same as K but "Pois" is straightened to a straight line
l_fun <- Lest(sp_sett)
plot(l_fun, xlim=c(0,5000), main="L-Function analysis")

# Pair Correlation Function analysis (PCF)
# Same principle as K, but different process ("donut rings")
# Expected Poisson is the green horizontal line
pc_fun <- pcf(sp_sett)
plot(pc_fun, xlim=c(0,5000), main="Pair Correlation Function analysis")

# Monte Carlo Simulation
pc_fun_100 <- envelope(sp_sett, pcf, nsim=99)
plot(pc_fun_100, xlim=c(0,20000), main="South: PCF with 99 MC Simulations")

```

## Task 2

Using Dataset 2, 
a) plot the distribution of medieval object findspots in the county of
Kent. 
b) Perform kernel density analysis on the findspots, using a few different search radii. 
c) Also perform a relative risk surface analysis, using the object type “seal matrix”.

```{r}


```

## Task 3

a) Perform aoristic analysis on Dataset 3, both on the whole dataset and on just the seal matrices. 
b) Include the graphs for aoristic weighing (bar charts) and 
c) for aoristic analysis with MC simulation (your choice of line, polygon “blocks colour” or boxplot charts), both for the overall and the seal matrices analyses.

```{r}


```

## Task 4

Perform PCA on dataset 4.

```{r}


```

## Task 5

Perform cluster analysis on dataset 4.

```{r}


```