---
title: "Week 1"
author: "Uine Kailamäki"
date: "2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## IMPORTING DATA IN R

In importing the .csv data to R there was a problem with importing European style .csv in which data is separated with ; instead of comma, and decimals are displayed with commas instead of dots. This was resolved by adding [sep=";", dec=","] when importing the data. I did also need to define the margin size for plots in order to display the results, but otherwise the exercise didn't pose much difficulty.

```{r, echo=FALSE}

# I tested if the file is readable and if my path is correct:

file.exists("~/R/SADE/week1/axes.csv")

# Importing the data

setwd("~/R/SADE/week1")
axes <- read.csv(file="axes.csv", header=TRUE, sep=";", dec=",")
class(axes)
head(axes)
summary(axes)
summary(axes$weight)
summary(axes$length)
sd(axes$weight)
sd(axes$thickness, na.rm=T)

# Plotting length vs weight

par(mai=c(1,1,1,1))
plot(x=axes$length, y=axes$weight, col="black", pch=16, main="Axe heads recovered by members of the public")

# More plots and histograms

boxplot(axes$length)

boxplot(axes$length, axes$width, ylab="millimeters")

barplot(axes$weight, ylab="grams")

barplot(sort(axes$weight), ylab="grams")

hist(axes$weight)

hist(axes$weight, breaks=seq(0, 1200, 100))

dev.new(device=pdf, height=6, width=12)
par(mfrow=c(1,2), mai=c(1,1,1,1))
barplot(sort(axes$weight), ylab="grams", names.arg =
c(axes$km_number), las=3)
hist(axes$weight)


```

## PLOTTING MAPS AND SPATIAL DATA

Question: GIT?
You can also embed plots, for example:

```{r, echo=FALSE}

library(rgdal)
library(raster)

setwd("~/R/SADE/week1")
fin <- readOGR(dsn="finland", layer="finland_valtakunta_simpler")

summary(fin)
plot(fin, col="lightgrey")
monu <- read.csv(file="monuments_fha.csv", header=TRUE, sep=",")
coordinates(monu) <- ~X+Y
crs(monu) <- crs(proj4string(fin))

points(monu, pch=19, cex=0.1)

points(monu[monu$period == "stone age", ], pch=19, cex=0.1, col="red")


# save.image("week1_sade.RData")
# load("week1_sade.RData")

```

